---
title: "PSI Data Analysis primer"
output: html_notebook
---

## Background information

In order to process large phenotyping data set, it is better to streamline this process as much as possible. In order to provide a primer for the less-experienced in R, and those who don't know where to start, I wrote this small example of data analysis and visualization. 

Remember - as the data that you are about to process is rather large - it is much better to document each step of your analysis to make sure it is as reproducible as possible. In the example data set I am using a collection of various mutants in Arabidopsis that were grown under "control" and "salt stress" conditions.

## Loading & organizing the datasets

OK - the data that is coming from the PSI machine is usually coming in separate .csv files > in order to make life easier for us we are going to combine all the phenotype containing files into one dataset:

```{r}
# load the datasets:
morpho_data <- read.csv("Rgb_Morpho_Plant.csv")
head(morpho_data)
```

In the example morpho_data:
- the treatment is indicated in collumn "Tray.Info"
- the genotype is indicated in collumns "Plant.Name"
- the individual plant identifier is indicated in collumn "Plant.ID"
- the collumns from "AREA_PX" to "SOL" represent the morphometric traits derived from the black-and-white image

Next we have data from other measuring stations as well:

```{r}
# data from fluorescent camera:
fc_data <- read.csv("Fc_Plant.csv")
# data from thermal/infrared camera:
ir_data <- read.csv("Ir_Plant.csv")
# data from RGB camera with color segmentation:
# WARNING - this data used to be not exported correctly - because the names of the HUEs was also containing semicolons - and required some manual manipulations of the table before importing it into R. 
col_data <- read.csv("Rgb_Color_Plant.csv")
# data from scales measuring the weight of each pot:
sc_data <- read.csv("ScalesMeasure.csv")
```

Let's combine all the datasets (except scales) together:

```{r}
library(reshape2)

# We will use columns "Tray.Info", "Round.Order", "Plant.ID" to make sure we are matching the right columns between individual datasets.

# In order to avoid unneccessary columns - we are going to first trim the datasets:
col_data <- col_data[,c(1:2,4:9,13:21)]
morpho_data <- morpho_data[,c(4:5,7,13:23)]
fc_data <- fc_data[,c(4:5,7,13:95)]
ir_data <- ir_data[,c(4:5,7,13:17)]

all_data <- merge(col_data, morpho_data, by=c("Tray.ID", "Round.Order", "Plant.ID"))
all_data <- merge(all_data, fc_data, by=c("Tray.ID", "Round.Order", "Plant.ID"))
# Then I have a little problem - because my IR is measured in a round before the full cycle - and the round order numbers are different. So for me to be able to merge two dataset I need to create new column "Round.Order2"
ir_data$Round.Order2 <- ir_data$Round.Order+1
# and then let's replace the original Round.Order by this tricky round order and remove it from the dataset:
ir_data$Round.Order <- ir_data$Round.Order2
ir_data <- ir_data[,1:8]
# and now we can fuse all together:
all_data <- merge(all_data, ir_data, by=c("Tray.ID", "Round.Order", "Plant.ID"))
head(all_data)
```

Just because I am obnoxious - we can save this file in the working directory as well

```{r}
write.csv(all_data, "pheno_data.csv", row.names = F)
```

## 


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

This notebook was created by [Magdalena Julkowska](https://mmjulkowska.github.io) (KAUST, Saudi Arabia), a PostDoc at Mark Tester group. Released under MIT license. 